Non-Vertical Cultural Transmission, Assortment, \\and the Evolution of Cooperation

### 2021-04-26 ms/main.tex
move SM title to correct place
Fix typo and remove todo
use single pdf file figures

% TODO change "scoail association" to "interaction-transmission association"

@@

\section*{Figures}
%%% Figure: model illustration
\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{Illustration.pdf}
  \caption{\textbf{Model illustration.} 
  \textbf{(a)}~First, offspring inherit their parent's phenotype via vertical cultural transmission with probability $v$, or the phenotype of a random non-parental adult via oblique cultural transmission with probability $1-v$.
  \textbf{(b)}~Second, adults socially interact in pairs in a prisoner's dilemma game. Horizontal cultural transmission occurs from a random individual in the population, with probability $1-\alpha$, or from the social partner, with probability $\alpha$, where $\alpha$ is the interaction-transmission association parameter.
  \textbf{(c)}~The prisoner's dilemma payoff matrix shows the fitness of phenotype $\phi_1$ when interacting with phenotype $\phi_2$.
    \textbf{(d)}~The probabilities of successful horizontal cultural transmission of phenotypes $A$ (cooperator) and $B$ (defector) are $T_A$ and $T_B$, respectively.
    }
  \label{fig:illustration}
\end{figure}

@@

\begin{figure}[p]
  \centering       
    \includegraphics[width=0.45\textwidth]{Result3_TA_TB.eps}~
    \includegraphics[width=0.45\textwidth]{Result3_TB_TA.eps}
    \includegraphics[width=0.45\textwidth]{Result3_alpha_Vs_v_TA_TB.eps}~
    \includegraphics[width=0.45\textwidth]{Result3_alpha_Vs_v_TB_TA.eps}
    \caption{\textbf{Evolution of cooperation under vertical, oblique, and horizontal cultural transmission.} 
    The figure shows parameter ranges for global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable polymorphism (blue), and stable polymorphism of cooperation and defection (yellow).
	In all cases the vertical transmission rate $v$ is on the x-axis.
	(\textbf{a-b})~Cost of cooperation $c$ is on the y-axis and the cost thresholds $\gamma_1$ and $\gamma_2$ (Eqs.~\ref{eq:cost_thresholds_v_threshold}) are represented by the solid and dashed lines, respectively. 
    (\textbf{c-d})~Interaction-transmission association $\alpha$ is on the y-axis and the interaction-transmission association thresholds $a_1$ and $a_2$ (Eqs.~\ref{eq:boundries_assortative_meeting_general_case}) are represented by the solid and dashed lines, respectively. 
    Horizontal transmission is biased in favor of cooperation, $T_A>T_B$, in (\textbf{a}) and (\textbf{c}), or defection, $T_A<T_B$, in (\textbf{b}) and (\textbf{d}).    
    Here, $T_A = 0.5$, and
    (\textbf{a}) $b=1.2$, , $T_B = 0.4$, $\alpha = 0.4$;
    (\textbf{b}) $b=2$, $T_B = 0.7$, $\alpha = 0.7$;
    (\textbf{c}) $b=1.2$, $T_B = 0.4$, $c=0.5$;
    (\textbf{d}) $b=2$, $T_B = 0.7$, $c=0.5$.
    }
    \label{fig:equilibria}
    %TODO update y label on c d
\end{figure}

@@

\begin{figure}[p]
  \centering       
    \includegraphics[width=0.45\textwidth]{Result2_c_zero_alpha.eps}
	  ~
    \includegraphics[width=0.45\textwidth]{Result2_c_non_zero_alpha.eps}
    ~
    \includegraphics[width=0.45\textwidth]{Result2_alpha.eps}   
    ~
    \includegraphics[width=0.48\textwidth]{coexistence_without_oblique.eps}
    \caption{\textbf{Evolution of cooperation under vertical and horizontal cultural transmission (v=1).} 
    The figure shows parameter ranges for global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable polymorphism (blue), and stable polymorphism of cooperation and defection (yellow).
    (\textbf{a-c})~The horizontal transmission bias ($T_A-T_B$) is on the x-axis.
    In panels \textbf{(a)} and \textbf{(b)}, the cost of cooperation $c$ is on the y-axis and the cost thresholds $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_thresholds_v_threshold}) are the solid and dashed lines, respectively. 
	In panel \textbf{(c)}, interaction-transmission association $\alpha$ is on the y-axis and the interaction-transmission association thresholds $a_1$ and $a_2$ (Eqs.~\ref{eq:boundries_assortative_meeting_general_case}) are the solid and dashed lines, respectively.
    Here, $b=1.3$, $T_A=0.4$, $v=1$, (a)~$\alpha = 0$, (b)~$\alpha = 0.7$, (c)~$c = 0.35$.    
    \textbf{(d)}~Change in frequency of cooperation among juveniles ($\hat{p}'-\hat{p}$) as a function of the frequency ($\hat{p}$), see \autoref{eq:nextgen_juveniles}.
    The orange curve shows convergence to a stable polymorphism ($T_A = 0.4$, $T_B = 0.9$, $b = 12$, $c=0.35$, $v=1$, and $\alpha = 0.45$). % todo ref eq 12
    The blue curve shows fixation of either cooperation or defection, depending on the initial frequency ($T_A = 0.5$, $T_B = 0.1$, $b = 1.3$, $c=0.904$, $v=1$, and $\alpha = 0.4$).
    Black circles show the three equilibria.
  	}
    \label{fig:equilibria_v1}
    %TODO update y label on c d
\end{figure}

%%% Figure: Spatial model - global selection
\begin{figure}[p]
  \centering
	\includegraphics[width=0.3\textwidth]{{Ta=Tb_b=1.3_globalSelection_a}.pdf}
    ~
	\includegraphics[width=0.3\textwidth]{{Ta=0.4_Tb=0.3-0.5_b=1.3_globalSelection_withSmallBox_b}.pdf} 
    ~
    \includegraphics[width=0.36\textwidth]{{timeSereis_b=1.3_TA=0.4_TB=0.435_c=0.02_globalSelection_normalizeByNumInteractions_c}.pdf}
  \caption{
  \textbf{Evolution of cooperation in a structured population.}
	\textbf{(a-b)}~The expected frequency of cooperators in a structured population after 10,000 generations is shown (red for 0\%, green for 100\%) as a function of both the cost of cooperation, $c$, on the y-axis, and either the symmetric horizontal transmission rate, $T=T_A=T_B$, on the x-axis of panel~\textbf{(a)}, or the transmission bias, $T_A-T_B$, on the x-axis of panel~\textbf{(b)}.
  Black curves represent the cost thresholds for the evolution of cooperation in a well-mixed population with interaction-transmission association, where $\alpha=1/8$ in inequality~\ref{eq:equal_transmission} for  panel~\textbf{(a)} and in Eqs.~\ref{eq:cost_thresholds_v_threshold} for panel~\textbf{(b)}.
	The inset in panel~\textbf{(b)} focuses on an area of the parameter range in which neither phenotype is fixed throughout the simulation, maintaining a stochastic locally stable polymorphism~\citep{Karlin1975}.
	This stochastic polymorphism is illustrated in panel~\textbf{(c)}, which shows the frequency of cooperators (green) and defectors (red) over time for the parameter set marked by an~\textit{x} in panel~\textbf{(b)}.
  In all cases, the population evolves on a 100-by-100 grid.
  Cooperation and horizontal transmission are both local between neighbouring sites, and each site has 8 neighbours.
  Selection operates globally (see \autoref{fig:spatial_local_selection} for results from a model with local selection).
  Simulations were stopped at generation 10,000 or if one of the phenotypes fixed. 50 simulations were executed for each parameter set.
  Benefit of cooperation, $b=1.3$; perfect vertical transmission $v=1$.
  \textbf{(a)} Symmetric horizontal transmission, $T=T_A=T_B$;
  \textbf{(b)} Horizontal transmission rate $T_A$ is fixed at $0.4$, and $T_B$ varies, $0.3<T_B<0.5$.
  \textbf{(c)} Horizontal transmission rates $T_A = 0.4 < T_B = 0.435$ and cost of cooperation $c=0.02$.
  }
  \label{fig:spatial}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section*{\huge{Supplementary material}}
\beginsupplement % https://support.authorea.com/en-us/article/how-to-create-an-appendix-section-or-supplementary-information-1g25i5a/

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendices
\begin{appendices}
\renewcommand{\theequation}{\thesection\arabic{equation}}
\counterwithin*{equation}{section}

@@

% Supporting figures
\newpage
\section*{Supporting figures}

%%% Figure: Invasion
\begin{figure}[h]
  \centering
   \includegraphics[scale=0.5]{Figures/invasion.eps}
  \caption{
  \textbf{Reduction principle for interaction-transmission association.} 
  Consecutive fixation of modifier alleles that reduce interaction-transmission association $\alpha$ in numerical simulations of evolution with two modifier alleles (\autoref{eq:next_gen_p_1}).
  When an invading modifier allele is established in the population (frequency > 99.95\%), a new modifier allele that reduces interaction-transmission association by 5\% is introduced (at initial frequency 0.5\%).
  \textbf{(a)}~The frequency of the cooperative phenotype $A$ over time.
  \textbf{(b)}~The frequency of the invading modifier allele $m$ over time. 
  \textbf{(c)}~The population mean fitness ($\bar{w}$) over time.
  Here, $c = 0.05$, $b=1.3$, $T_A=0.4<T_B=0.7$, initial interaction-transmission association $\alpha_1=0.7$, lower interaction-transmission association threshold $a_2=0.605$.  
  }
  \label{fig:invasion}
\end{figure}

%%% Figure: Spatial model - local selection
\begin{figure}[h]
  \centering
   \includegraphics[scale=0.5]{{Ta=Tb_b=1.3_localSelection}.pdf}
    ~
    \includegraphics[scale=0.5]{{Ta=0.4_Tb=0.3-0.5_b=1.3_localSelection}.pdf} \\
  \caption{
  \textbf{Evolution of cooperation in a structured population with local selection.}
  The expected frequency of cooperators in a structured population after 10,000 generations is shown (red for 0\%, green for 100\%) as a function of both the cost of cooperation ($c$) on the y-axis, and the symmetric horizontal transmission rate ($T=T_A=T_B$) on the x-axis of panel~\textbf{(a)}, or the transmission bias $T_A-T_B$ on the x-axis of panel~\textbf{(b)}.
  Cooperation and horizontal transmission are both local between neighbouring sites, and each site had 8 neighbours.
  Selection operates locally (see \autoref{fig:spatial} for results from a model with global selection).
  The black curves represent the cost thresholds for the evolution of cooperation in a well-mixed population with interaction-transmission association, where $\alpha=1/8$ in inequality~\ref{eq:equal_transmission} for panel~\textbf{(a)} and in Eqs.~\ref{eq:cost_thresholds_v_threshold} for panel~\textbf{(b)}.
  The population evolves on a 100-by-100 grid.
  Simulations were stopped at generation 10,000 or if one of the phenotypes fixed.
  50 simulations were executed for each parameter set.
  Here, benefit of cooperation, $b=1.3$; perfect vertical transmission $v=1$.
  \textbf{(a)} Symmetric horizontal transmission, $T=T_A=T_B$.
  \textbf{(b)} Horizontal transmission rate $T_A$ is fixed at $0.4$, and $T_B$ varies, $0.3<T_B<0.5$.
  }
  \label{fig:spatial_local_selection}
\end{figure}

>>>>

@@

\section*{Figures}
%%% Figure: model illustration
\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth]{PRSB_figures/fig_1.pdf}
  \caption{\textbf{Model illustration.} 
  \textbf{(a)}~First, offspring inherit their parent's phenotype via vertical cultural transmission with probability $v$, or the phenotype of a random non-parental adult via oblique cultural transmission with probability $1-v$.
  \textbf{(b)}~Second, adults socially interact in pairs in a prisoner's dilemma game. Horizontal cultural transmission occurs from a random individual in the population, with probability $1-\alpha$, or from the social partner, with probability $\alpha$, where $\alpha$ is the interaction-transmission association parameter.
  \textbf{(c)}~The prisoner's dilemma payoff matrix shows the fitness of phenotype $\phi_1$ when interacting with phenotype $\phi_2$.
    \textbf{(d)}~The probabilities of successful horizontal cultural transmission of phenotypes $A$ (cooperator) and $B$ (defector) are $T_A$ and $T_B$, respectively.
    }
  \label{fig:illustration}
\end{figure}

@@

\begin{figure}[p]
  \centering       
    \includegraphics[width=1\textwidth]{PRSB_figures/fig_2.pdf}
    \caption{\textbf{Evolution of cooperation under vertical, oblique, and horizontal cultural transmission.} 
    The figure shows parameter ranges for global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable polymorphism (blue), and stable polymorphism of cooperation and defection (yellow).
	In all cases the vertical transmission rate $v$ is on the x-axis.
	(\textbf{a-b})~Cost of cooperation $c$ is on the y-axis and the cost thresholds $\gamma_1$ and $\gamma_2$ (Eqs.~\ref{eq:cost_thresholds_v_threshold}) are represented by the solid and dashed lines, respectively. 
    (\textbf{c-d})~Interaction-transmission association $\alpha$ is on the y-axis and the interaction-transmission association thresholds $a_1$ and $a_2$ (Eqs.~\ref{eq:boundries_assortative_meeting_general_case}) are represented by the solid and dashed lines, respectively. 
    Horizontal transmission is biased in favor of cooperation, $T_A>T_B$, in (\textbf{a}) and (\textbf{c}), or defection, $T_A<T_B$, in (\textbf{b}) and (\textbf{d}).    
    Here, $T_A = 0.5$, and
    (\textbf{a}) $b=1.2$, , $T_B = 0.4$, $\alpha = 0.4$;
    (\textbf{b}) $b=2$, $T_B = 0.7$, $\alpha = 0.7$;
    (\textbf{c}) $b=1.2$, $T_B = 0.4$, $c=0.5$;
    (\textbf{d}) $b=2$, $T_B = 0.7$, $c=0.5$.
    }
    \label{fig:equilibria}
\end{figure}

@@

\begin{figure}[p]
  \centering       
	\includegraphics[width=1\textwidth]{PRSB_figures/fig_3.pdf}
    \caption{\textbf{Evolution of cooperation under vertical and horizontal cultural transmission (v=1).} 
    The figure shows parameter ranges for global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable polymorphism (blue), and stable polymorphism of cooperation and defection (yellow).
    (\textbf{a-c})~The horizontal transmission bias ($T_A-T_B$) is on the x-axis.
    In panels \textbf{(a)} and \textbf{(b)}, the cost of cooperation $c$ is on the y-axis and the cost thresholds $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_thresholds_v_threshold}) are the solid and dashed lines, respectively. 
	In panel \textbf{(c)}, interaction-transmission association $\alpha$ is on the y-axis and the interaction-transmission association thresholds $a_1$ and $a_2$ (Eqs.~\ref{eq:boundries_assortative_meeting_general_case}) are the solid and dashed lines, respectively.
    Here, $b=1.3$, $T_A=0.4$, $v=1$, (a)~$\alpha = 0$, (b)~$\alpha = 0.7$, (c)~$c = 0.35$.    
    \textbf{(d)}~Change in frequency of cooperation among juveniles ($\hat{p}'-\hat{p}$) as a function of the frequency ($\hat{p}$), see \autoref{eq:nextgen_juveniles}.
    The orange curve shows convergence to a stable polymorphism ($T_A = 0.4$, $T_B = 0.9$, $b = 12$, $c=0.35$, $v=1$, and $\alpha = 0.45$). % todo ref eq 12
    The blue curve shows fixation of either cooperation or defection, depending on the initial frequency ($T_A = 0.5$, $T_B = 0.1$, $b = 1.3$, $c=0.904$, $v=1$, and $\alpha = 0.4$).
    Black circles show the three equilibria.
  	}
    \label{fig:equilibria_v1}
\end{figure}

%%% Figure: Spatial model - global selection
\begin{figure}[p]
  \centering
  \includegraphics[width=1\textwidth]{PRSB_figures/fig_4.pdf}
  \caption{
  \textbf{Evolution of cooperation in a structured population.}
	\textbf{(a-b)}~The expected frequency of cooperators in a structured population after 10,000 generations is shown (red for 0\%, green for 100\%) as a function of both the cost of cooperation, $c$, on the y-axis, and either the symmetric horizontal transmission rate, $T=T_A=T_B$, on the x-axis of panel~\textbf{(a)}, or the transmission bias, $T_A-T_B$, on the x-axis of panel~\textbf{(b)}.
  Black curves represent the cost thresholds for the evolution of cooperation in a well-mixed population with interaction-transmission association, where $\alpha=1/8$ in inequality~\ref{eq:equal_transmission} for  panel~\textbf{(a)} and in Eqs.~\ref{eq:cost_thresholds_v_threshold} for panel~\textbf{(b)}.
	The inset in panel~\textbf{(b)} focuses on an area of the parameter range in which neither phenotype is fixed throughout the simulation, maintaining a stochastic locally stable polymorphism~\citep{Karlin1975}.
	This stochastic polymorphism is illustrated in panel~\textbf{(c)}, which shows the frequency of cooperators (green) and defectors (red) over time for the parameter set marked by an~\textit{x} in panel~\textbf{(b)}.
  In all cases, the population evolves on a 100-by-100 grid.
  Cooperation and horizontal transmission are both local between neighbouring sites, and each site has 8 neighbours.
  Selection operates globally (see \autoref{fig:spatial_local_selection} for results from a model with local selection).
  Simulations were stopped at generation 10,000 or if one of the phenotypes fixed. 50 simulations were executed for each parameter set.
  Benefit of cooperation, $b=1.3$; perfect vertical transmission $v=1$.
  \textbf{(a)} Symmetric horizontal transmission, $T=T_A=T_B$;
  \textbf{(b)} Horizontal transmission rate $T_A$ is fixed at $0.4$, and $T_B$ varies, $0.3<T_B<0.5$.
  \textbf{(c)} Horizontal transmission rates $T_A = 0.4 < T_B = 0.435$ and cost of cooperation $c=0.02$.
  }
  \label{fig:spatial}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

% Appendices
\begin{appendices}
\renewcommand{\theequation}{\thesection\arabic{equation}}
\counterwithin*{equation}{section}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\section*{\huge{Supplementary material}}
\beginsupplement % https://support.authorea.com/en-us/article/how-to-create-an-appendix-section-or-supplementary-information-1g25i5a/

%%% Figure: Invasion
\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth]{PRSB_figures/fig_s1.pdf}
  \caption{
  \textbf{Reduction principle for interaction-transmission association.} 
  Consecutive fixation of modifier alleles that reduce interaction-transmission association $\alpha$ in numerical simulations of evolution with two modifier alleles (\autoref{eq:next_gen_p_1}).
  When an invading modifier allele is established in the population (frequency > 99.95\%), a new modifier allele that reduces interaction-transmission association by 5\% is introduced (at initial frequency 0.5\%).
  \textbf{(a)}~The frequency of the cooperative phenotype $A$ over time.
  \textbf{(b)}~The frequency of the invading modifier allele $m$ over time. 
  \textbf{(c)}~The population mean fitness ($\bar{w}$) over time.
  Here, $c = 0.05$, $b=1.3$, $T_A=0.4<T_B=0.7$, initial interaction-transmission association $\alpha_1=0.7$, lower interaction-transmission association threshold $a_2=0.605$.  
  }
  \label{fig:invasion}
\end{figure}

%%% Figure: Spatial model - local selection
\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth]{PRSB_figures/fig_s2.pdf}
  \caption{
  \textbf{Evolution of cooperation in a structured population with local selection.}
  The expected frequency of cooperators in a structured population after 10,000 generations is shown (red for 0\%, green for 100\%) as a function of both the cost of cooperation ($c$) on the y-axis, and the symmetric horizontal transmission rate ($T=T_A=T_B$) on the x-axis of panel~\textbf{(a)}, or the transmission bias $T_A-T_B$ on the x-axis of panel~\textbf{(b)}.
  Cooperation and horizontal transmission are both local between neighbouring sites, and each site had 8 neighbours.
  Selection operates locally (see \autoref{fig:spatial} for results from a model with global selection).
  The black curves represent the cost thresholds for the evolution of cooperation in a well-mixed population with interaction-transmission association, where $\alpha=1/8$ in inequality~\ref{eq:equal_transmission} for panel~\textbf{(a)} and in Eqs.~\ref{eq:cost_thresholds_v_threshold} for panel~\textbf{(b)}.
  The population evolves on a 100-by-100 grid.
  Simulations were stopped at generation 10,000 or if one of the phenotypes fixed.
  50 simulations were executed for each parameter set.
  Here, benefit of cooperation, $b=1.3$; perfect vertical transmission $v=1$.
  \textbf{(a)} Symmetric horizontal transmission, $T=T_A=T_B$.
  \textbf{(b)} Horizontal transmission rate $T_A$ is fixed at $0.4$, and $T_B$ varies, $0.3<T_B<0.5$.
  }
  \label{fig:spatial_local_selection}
\end{figure}
